\LARGE \bf The Battle for Filter Supremacy: A Retrospective: The Movie: The Game: The Paper

### 2014-12-29 Report/project_report.tex
Missed a few primes
Cleaned up the SWF notation

In this work, we compare and contrast three modern solutions, the Extended Kalman Filter (EKF), the Sliding Window Filter (SWF), and a hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.

@@

[EKF EQUATIONS HERE?]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.

\begin{equation}
(\Vec H^T \Vec T ^{-1} \Vec H ) \delta \Vec{x}^*  = - \Vec{H}^T\Vec{T}^{-1}\Vec{e}(\bar{\Vec{x}})
\end{equation}

\begin{equation}
\delta\Vec{x} := \bbm \delta \Vec x_0 \\ \vdots \\ \delta \Vec x_K \\ \delta \Vec f_1 \\ \vdots \\ \delta \Vec f_N\ebm,
\end{equation}
where $N$ is the number of unique features observed in the window and 
\begin{equation}
\mathbf{H} := \left[\begin{array}{cccccc}
 -\Vec H_{\Vec x, 1} & \Vec 1 & &  & &   \\ 
 & -\Vec G_{\Vec x, 1} &  &  &  &   -\Vec G_{\Vec f, 1} \\ \hline
 & -\Vec H_{\Vec x, 2} & \Vec 1 & & &    \\ 
 & & -\Vec G_{\Vec x, 2} &  & &  -\Vec G_{\Vec f, 2}   \\ \hline
 &  & \ddots & \ddots & & \vdots  \\  \hline
  & & & -\Vec H_{\Vec x, K} & \Vec 1 & \\ 
 & & & & -\Vec G_{\Vec x, K} &  -\Vec G_{\Vec f, K}   \\  
\end{array}\right], 
\end{equation},

\begin{equation}
\Vec G_{\Vec{x}, k}^j := \pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} \bbm -\Vec C_{cv} \Vec C_{v_k i} & \Vec C_{cv} (\Vec C_{v_k i} (\boldsymbol{\rho}_i^{p_j i} - \Vec r_i^{v_k i} ) )^\times  \ebm\end{equation}

\begin{equation}
\quad \Vec G_{\Vec x, k} := \bbm \Vec G_{\Vec x, k}^1 \\ \vdots \\ \Vec G_{\Vec x, k}^M \ebm
\end{equation}

\begin{equation}
\Vec G_{\Vec f, k}^j := \pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} \bbm \Vec C_{cv} \Vec C_{v_k i}\ebm\end{equation}

\begin{equation}
\quad \Vec G_{\Vec f, k} := \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature id $n$}} \in \Real^{4M \times 3N}
\end{equation}

and,
\begin{equation}
\bar{\Vec p}_{c_k}^{p_j c_k} := \bbm x \\ y \\ z \ebm = \Vec C_{cv} \left( \Vec C_{v_k i} (\boldsymbol{\rho}_i^{p_j i} - \Vec r_i^{v_k i} )   - \boldsymbol \rho_v^{cv} \right),
\end{equation}

\begin{equation}
\pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y  \\ \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u (x-b) \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y\ebm.
\end{equation}
or for monocular,
\begin{equation}
\pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y \ebm.
\end{equation}

@@

We obtain the predicted IMU-IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
\begin{align} \label{eq:imucovar}
    \dot{\hat{\Vec{P}}}_{II}\left(t_k, t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \notag \\ 
                & + \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \Vec{F}^T \notag \\ 
                & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
\end{align}
with the initial condition $\hat{\Vec{P}}_{II}\left(t_k, t_k\right) = \hat{\Vec{P}}_{II,k}$.

@@

Next, we define a vector of three parameters:
\begin{align}
\hat{\Vec{y}} &= \bbm \alpha & \beta & \gamma \ebm^T = \frac{1}{\hat{Z}^{(j)}_{C_1}} \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & 1 \ebm ^T
\end{align}
The feature position in camera frame $C_i$ can be expressed as
\begin{align}
\hat{\Vec{p}}_{C_i}^{f_j C_i} = \hat{\Vec{C}}_{i1} \hat{\Vec{p}}_{C_1}^{f_j C_1}  + \hat{\Vec{p}}_{C_i}^{C_1 C_i}
\end{align}
With our state parameters, we can use this equation to define three functions:
\begin{align}
    \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \\ h_3 (\hat{\Vec{y}}) \ebm 
    & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i} \notag \\
    & = \bbm C_{11} \alpha + C_{12} \beta + C_{13} + \gamma \hat{x}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{21} \alpha + C_{22} \beta + C_{23} + \gamma \hat{y}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
        \ebm 
\end{align}
The camera measurement error can then be written as
\begin{align}
\Vec e(\hat{\Vec{y}}) = \hat{\Vec{z}}_i^{(j)} - \frac{1}{h_3 (\hat{\Vec{y}}) } \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \ebm
\end{align}
The least squares linear system then becomes
\begin{align}
(\Vec  E^T \Vec W ^{-1} \Vec E ) \delta \Vec{y}^*  = - \Vec{E}^T\Vec{W}^{-1}\Vec{e}(\hat{\Vec{y}})
\end{align}
where
\begin{align}
\Vec E &= \frac{\partial \Vec e}{\partial \hat{\Vec{y}}}
& &
\text{and}
&
\Vec W &= \bbm \Vec{R}_1^{(j)} & & \\ & \ddots & \\ & & \Vec{R}_N^{(j)} \ebm 
\end{align}
with $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_u, \sigma^2_v\right\}$.

>>>>

In this work, we compare and contrast three modern solutions, the Extended Kalman Filter (EKF), the Sliding Window Filter (SWF), and a hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
We present an experimental comparison of the SWF and MSCKF using a dataset consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
\begin{equation}
(\Vec H^T \Vec T ^{-1} \Vec H ) \delta \Vec{x}^*  = - \Vec{H}^T\Vec{T}^{-1}\Vec{e}(\bar{\Vec{x}})
\end{equation}
\begin{equation}
\delta\Vec{x} := \bbm \delta \Vec x_0 & \hdots & \delta \Vec x_K & \delta \Vec f_1 & \hdots & \delta \Vec f_N\ebm ^T,
\end{equation}
where $N$ is the number of unique features observed in the window.
The matrix $\Vec{H}$ is given by (showing non-zero blocks only)
\begin{equation}
\mathbf{H} = \left[\begin{array}{cccccc}
 -\Vec H_{\Vec x, 1} & \Vec 1 & &  & &   \\ 
 & -\Vec G_{\Vec x, 1} &  &  &  &   -\Vec G_{\Vec f, 1} \\ \hline
 & -\Vec H_{\Vec x, 2} & \Vec 1 & & &    \\ 
 & & -\Vec G_{\Vec x, 2} &  & &  -\Vec G_{\Vec f, 2}   \\ \hline
 &  & \ddots & \ddots & & \vdots  \\  \hline
  & & & -\Vec H_{\Vec x, K} & \Vec 1 & \\ 
 & & & & -\Vec G_{\Vec x, K} &  -\Vec G_{\Vec f, K}   \\  
\end{array}\right], 
\end{equation}
where
\begin{align}
\Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\boldsymbol{\rho}_i^{f_j G} - \Vec r_G^{IG,k} ) )^\times  \ebm ^T
\\[0.5em]
\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
\\[0.5em]
\Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k}
\\[0.5em]
\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j$}}
\end{align}
with
\begin{equation}
\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\boldsymbol{\rho}_G^{f_j G} - \Vec r_G^{IG,k} )   - \boldsymbol \rho_I^{CI} \right),
\end{equation}

\begin{equation}
\pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} = \frac{1}{z^2} \bbm f_u z & 0 & - f_u x \\  0 & f_v z & -f_v y \ebm.
\end{equation}

@@

We obtain the predicted IMU-IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
\begin{align} \label{eq:imucovar}
    \dot{\hat{\Vec{P}}}_{II}\left(t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) 
                \hat{x}_{C_i}^{C_1 C_i} + \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) \Vec{F}^T \notag \\ 
                & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
\end{align}
with the initial condition $\hat{\Vec{P}}_{II}\left(t_k\right) = \hat{\Vec{P}}_{II,k}$.

@@

Next, we define a vector of three parameters:
\begin{align}
\hat{\Vec{y}} &= \bbm \alpha & \beta & \gamma \ebm^T = \frac{1}{\hat{Z}^{(j)}_{C_1}} \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & 1 \ebm ^T
\end{align}
The feature position in camera frame $C_i$ can be expressed as
\begin{align}
\hat{\Vec{p}}_{C_i}^{f_j C_i} = \hat{\Vec{C}}_{i1} \hat{\Vec{p}}_{C_1}^{f_j C_1}  + \hat{\Vec{p}}_{C_i}^{C_1 C_i}
\end{align}
With our state parameters, we can use this equation to define three functions:
\begin{align}
    \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \\ h_3 (\hat{\Vec{y}}) \ebm 
    & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i} \notag \\
    & = \bbm C_{11} \alpha + C_{12} \beta + C_{13} + \gamma \hat{x}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{21} \alpha + C_{22} \beta + C_{23} + \gamma \hat{y}_{C_i}^{C_1 C_i} \\[0.5em]
            C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
        \ebm 
\end{align}
The camera measurement error can then be written as
\begin{align}
\Vec e(\hat{\Vec{y}}) = \hat{\Vec{z}}_i^{(j)} - \frac{1}{h_3 (\hat{\Vec{y}}) } \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \ebm
\end{align}
The least squares linear system then becomes
\begin{align}
(\Vec  E^T \Vec W ^{-1} \Vec E ) \delta \Vec{y}^*  = - \Vec{E}^T\Vec{W}^{-1}\Vec{e}(\hat{\Vec{y}})
\end{align}
where
\begin{align}
\Vec E &= \frac{\partial \Vec e}{\partial \hat{\Vec{y}}}
& &
\text{and}
&
\Vec W &= \bbm \Vec{R}_1'^{(j)} & & \\ & \ddots & \\ & & \Vec{R}_N'^{(j)} \ebm 
\end{align}
with $\Vec{R}_i'^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.
